Placing the spotted T Tauri star LkCa 4 on an HR diagram

### 2016-12-08 document/submission/ms.tex
Working on Issue #53-- on to the vsini discussion.

% Theory of starspots
% Jackson \& Jeffries 2014ab needs to go somewhere.

Starspots have recently been observed with exceptional photometric precision from monitoring surveys targeting exoplanet transits \citep[\emph{e.g}][]{harrison11,davenport15}.  The single band light curves from such surveys cannot separate the relative areas and temperatures of the emitting regions, and therefore cannot yield estimates for the filling factor of spots without making assumptions about the relative temperature contrast. Contemporaneous or near-contemporaneous panchromatic photometric monitoring \citep{petrov94,bouvier95,grankin07,cody14} can break the degeneracy between starspot filling factor and starspot temperature but these measurements are still limited to detecting large longitudinally asymmetric starspots.  Transiting planets or eclipsing binaries passing in front of starspots \citep{desert11} can break geometric degeneracies, providing estimates for the characteristic sizes and lifetimes of starspots or starspot groups, but the number of such transiting systems and demands of high photometric precision limit wider application.  Zeeman Doppler Imaging \citep[ZDI]{donati14} produces the latitudinal and longitudinal distribution of starspots in reconstructed brightness maps based on models for spectro-polarimetric line shapes, but would be unable to identify uniform distributions of small spots.  All of these techniques would tend to underestimate the starspot areal coverage on stars with an isotropic distribution of small starspots, or pole-on stars or stars with longitudinally symmetric bands of starspots.

Spectroscopic strategies have the power to measure both the starspot areal coverage and temperature in any geometric configuration.  For weak-lined T Tauri stars (WTTSs), SED modeling \citep{wolk96} indicated the presence of cool spots, while \citet{grankin98} and \citet{venuti15} estimated spot temperature and coverage from rotational modulation of optical colors.  The TiO bands in optical spectra have been used to measure spots on WTTSs \citep{petrov94}, stars in the Pleiades \citep{fang2016}, and evolved stars \citep[e.g.][]{neff95,oneal01,oneal04}.  Recent spectroscopic detections of spots on young stars have focused on TW Hya, an active accretor, and DQ Tau, a close binary in which both components are accreting \citep{debes13,bary14}.

@@

Several WTTS in the Taurus-Auriga star-forming region have large amplitudes in periodic light variations, achieving $0.4-0.8$ mag in the $V-$band: LkCa 4, V410 Tau, LkCa 7, V827 Tau, V830 Tau, and V819 Tau \citep{grankin08}.  In a sample that is not biased to variable stars, 
WTTSs observed in NGC 2264 show photometric variations similar to--but of lower average amplitude than-- LkCa 4 \citep{cody14,venuti15}.  Modeling of polychromatic photometric monitoring of the locus of WTTSs indicates a lower limit to the true coverage fraction $f_{\mathrm{eq}}\sim10-30\%$ and a difference $\teffa-\teffb\sim500$ K.

@@

%\todo{Verify that the final term possesses a $q$ and not a $q^2$ term.}

>>>>

Rotational modulation from starspots has been understood as the cause of cyclic variations in weak-lined T Tauri stars (WTTSs) and some classical T Tauri stars \citep{vrba86,herbst94}.  Spots have recently been observed with exceptional photometric precision from monitoring surveys targeting exoplanet transits \citep[\emph{e.g}][]{harrison11,davenport15}.  The single band light curves from such surveys cannot separate the relative areas and temperatures of the emitting regions, and therefore cannot yield estimates for the filling factor of spots without making assumptions about the relative temperature contrast. Contemporaneous or near-contemporaneous panchromatic photometric monitoring \citep{herbst94,petrov94,bouvier95,grankin07,cody14} can break the degeneracy between starspot filling factor and starspot temperature but these measurements are still limited to detecting large longitudinally asymmetric starspots.  Transiting planets or eclipsing binaries passing in front of starspots \citep{desert11} can break geometric degeneracies, providing estimates for the characteristic sizes and lifetimes of starspots or starspot groups, but the number of such transiting systems and demands of high photometric precision limit wider application.  Zeeman Doppler Imaging \citep[ZDI]{donati14} produces the latitudinal and longitudinal distribution of starspots in reconstructed brightness maps based on models for spectro-polarimetric line shapes, but would be unable to identify uniform distributions of small spots.  All of these techniques would tend to underestimate the starspot areal coverage on stars with an isotropic distribution of small starspots, or pole-on stars or stars with longitudinally symmetric bands of starspots.

Spectroscopic strategies have the power to measure both the starspot areal coverage and temperature in any geometric configuration.  For WTTSs, SED modeling \citep{wolk96} indicated the presence of cool spots, while \citet{grankin98} and \citet{venuti15} estimated spot temperature and coverage from rotational modulation of optical colors.  The TiO bands in optical spectra have been used to measure spots on WTTSs \citep{petrov94}, stars in the Pleiades \citep{fang2016}, and evolved stars \citep[e.g.][]{neff95,oneal01,oneal04}.  Recent spectroscopic detections of spots on young stars have focused on TW Hya, an active accretor, and DQ Tau, a close binary in which both components are accreting \citep{debes13,bary14}.

@@

Several WTTS in the Taurus-Auriga star-forming region have large amplitudes in periodic light variations, achieving $0.4-0.8$ mag in the $V-$band: LkCa 4, V410 Tau \citep{herbst89}, LkCa 7, V827 Tau, V830 Tau, and V819 Tau \citep{grankin08}.  In a sample that is not biased to variable stars, WTTSs observed in NGC 2264 show photometric variations similar to--but of lower average amplitude than-- LkCa 4 \citep{cody14,venuti15}.  Modeling of polychromatic photometric monitoring of the locus of WTTSs indicates a lower limit to the true coverage fraction $f_{\mathrm{eq}}\sim10-30\%$ and a difference $\teffa-\teffb\sim500$ K.

@@
